Spectroscopic Characterization of New Er$^{3+}$ Host Crystals for Quantum Network Applications

### 2020-04-21 Draft.tex
Finish MgO

Unfortunately, using bulk doping methods it is impossible to grow crystals with lower concentrations of \erbium than in this previous work \cite{Dibos2017}. The Thompson lab has thus turned to ion implantation to achieve lower doses of \erbium in new potential crystal hosts.\footnote{Ion implantation is a process by which ions are accelerated into the surface of a crystal, and become lodged in the structure of the crystal. More information on this process is given in Section } Last fall, it was shown that \erbium could be successfully implanted in \TiO[], a crystal with low nuclear spins, with narrow optical linewidths \cite{Phenicie2019}. \notetoself{Given that \TiO worked, why continue searching? Not super clear on this myself. Does it have to do with attaching photonic crystals?}

In the interests of discovering potentially more optimal host crystals, a large batch of crystals was implanted with \erbium ions. For all these implanted crystals, it was necessary to determine experimentally whether \erbium was visible in the crystal, and if so, to determine its spectral properties to begin understanding how \erbium ions in this crystal would perform as single photon emitters. Where are the emission and excitation lines? How bright and broad are they? What can we learn about how the \erbium ions are sitting in the crystal lattice? These are the questions that motivate this work.

@@

In general, the number of crystal field parameters is infinite. However, various symmetries of the crystal can eliminate many of the possible parameters. For example, if the rare earth ion sees a crystal environment with cubic symmetry, only four crystal field parameters can be nonzero, $B_{4}^{0}$, $B_{4}^{4}$, $B_{6}^{0}$, and $B_{6}^{4}$, and only two of these are actually independent. Thus the problem is a two-parameter one. Lea, Leask and Wolf show an efficient way of calculating the crystal field matrix for diagonalization, 
\begin{equation}\label{eq:7}
  H = B_{4}(O_{4}^{0}+ 5\cdot O_{4}^{4}) + B_{6}(O_{6}^{0}-21\cdot O_{6}^{4}),
\end{equation}
where $O_{4}^{0}$, $O_{4}^{4}$, $O_{6}^{0}$ and $O_{6}^{4}$ are $2J+1$ dimensional constant matrices (for their exact form see \cite{Lea1962}). The upshot of this for our purposes is that the crystal field environment that \erbium ions experience in MgO, a crystal with cubic symmetry, could potentially be determined precisely. In Section \ref{sec:mgo} we take up this task.

@@

\textbf{Sample preparation}: Implanted with $10^{19}$ dose \notetoself{find out units on this}. Unannealed.

\textbf{Structure}: Magnesium oxide (MgO) is a cubic lattice consisting of O$^{2-}$ ions and Mg$^{2+}$ ions held together by ionic bonding. There are four types of sites with non-trivial symmetries: the two substitutional sites have cubic symmetry, as does the site in the center of the cube, and the midpoints of the ionic bonds have $C_{4}$ symmetry in one plane and $D_{1}$ symmetry along the axis normal to this plane. Because the crystal must remain net neutral, there might be some restructuring of the surrounding environment to compensate the introduction of the additional positive charge from the implanted \erbium ion. This may disturb the symmetries of the crystal.

\textbf{Energy levels}: The spectrum for MgO is fairly complicated, with at least four clearly different sites contributing to the excitation-emission spectrum. The experimentally determined energy levels for each of these sites is listed in Table \ref{tab:mgolevels}. A few idiosyncracies are important to note for some of these sites. The $Z_{5}$ energy at Site 1 was never directly observed in the form of $Y_{1}Z_{5}$ emission like all the other $Z$ levels, and is hence listed in parenthesis. However, its location can be inferred from an emission line $Y_{3}Z_{5}$ which appears at higher temperatures. This peak can be seen emerging in Figure \ref{fig:mgosite1} and Figure \ref{fig:mgotemp}.

@@

A third idiosyncracy has to do with the brightest peak in the entire spectrum, which belongs to Site 3, and falls at 6540 \wn in excitation. This peak is bimodal, as shown in Figure \notetoself{make this figure}. There are two explanations for this distribution. Either these two peaks correspond to different levels in the same site, or they correspond to the same level at two sites which are very similar but slightly different.

Our energy level assignments are supported by the plots in Figure \ref{fig:mgotemp}. These figures show the emission spectra when exciting the $Z_{1}Y_{1}$ transition of the different sites, normalized so that emission at $Y_{1}Z_{1}$ is equal to 1. The other features identified as emission from $Y_{1}$ remain the same height relative to $Y_{1}Z_{1}$. The features identified as emission from higher $Y$ levels, on the other hand, get brighter as the temperature increases, demonstrating that their thermal occupation has increased. An important assumption in this situation is that emission is slow enough that the excited state electrons thermalize fully before decaying. This assumption is supported by the fact that the relative heights of the peaks in the emission spectrum are independent of the excitation transition at which they are observed.

\textbf{Crystal field modelling}: Unimplanted MgO has a cubic crystal structure. In an environment with cubic symmetry, there are only two free crystal field parameters, and thus the task of fitting an experimentally determined spectrum to theory is feasible (see Section \ref{sec:crystal-fields}). Interestingly, however, neither Site 1, Site 2, or Site 4 have energy levels consistent with a cubic crystal environment. From this, we conclude that charge compensation when \erbium replaces Mg$^{2+}$ necessitates changes to the crystal structure surrounding the \erbium which break the cubic symmetry of the environment. A likely candidate is the removal of nearby Mg$^{2+}$ ions, which can occur in line with the \erbium ion two sites away, or diagonally from the \erbium ion, or possibly farther away.

@@

\begin{table}[t]
  \centering
  \begin{tabular}{| c | c | c | c | c |}
    \hline
    MgO & \multicolumn{2}{c|}{Site 1}& \multicolumn{2}{c|}{Site 2} \\
    \hline
    $n$ & $Z_{n}$ & $Y_{n}$ ($+6499$ \wn) & $Z_{n}$ & $Y_{n}$ ($+6510$ \wn) \\
    \hline 
    1 & 0.0 & 0.0 & 0.0 & 0.0 \\
    2 & 26.7 & 7.8 & 41.2 & 29.1 \\
    3 & 95.3 & 34.9 & 83.8 & 58.3 \\
    4 & 128.6 & 71.7 & 156.4 & 117.2 \\
    5 & (195.2) & 131.3 & 187.2 & * \\
    6-8 & * & * & * & * \\
    \hline \hline
    MgO & \multicolumn{2}{c|}{Site 3}& \multicolumn{2}{c|}{Site 4} \\
    \hline
    $n$ & $Z_{n}$ & $Y_{n}$ ($+6492$ \wn) & $Z_{n}$ & $Y_{n}$ ($+6557$ \wn)\\
    \hline 
    1 & 0.0 & 0.0 & 0.0 & 0.0  \\
    2 & * & 48.1 & 82.8 & 42.0 \\
    3 & * & (48.4) & 168.6 & *  \\
    4 & * & (98.6) & * & * \\
    5-8 & * & * & * & * \\
    \hline
  \end{tabular}
  \caption{Experimentally determined energy levels of MgO for all visible sites. Transitions not observed are marked with an asterisk. The level $Z_{5}$ in Site 1 is not directly observed, and thus marked with parentheses. $Z_{2}$ and $Z_{3}$ for Site 3 may or not correspond to distinct levels, which would affect the numbering of $Z_{4}$; hence $Z_{3}$ and $Z_{4}$ are marked with parentheses.}
  \label{tab:mgolevels}
\end{table}

\textbf{Linewidths and shapes}: PLE scans of MgO were taken at 13, 32, 60, and 106K. Taking the level assignments discussed above, the excitation peaks corresponding to excitation from $Z_{1}$ to $Y_{n}$ for all visible $Y_{n}$ were located for all four sites. For each of these peaks, two profiles were fitted, a Gaussian profile and a Lorentzian profile. When the dominant source of broadening in a spectral line is homogeneous, its profile will be Lorentzian; when the dominant broadening is inhomogeneous, the profile will in general be Gaussian. Table \ref{tab:mgolinewidths} shows the results of these fits.

\begin{table}[t]
  \centering
  \begin{tabular}{|c| c| c | c | c | c | c | c | c|}
    \hline
    MgO & \multicolumn{4}{c|}{Site 1} & \multicolumn{4}{c|}{Site 2} \\
    \hline
    $n$ & 13K & 32K & 60K & 106K & 13K & 32K & 60K & 106K \\
    \hline
    1 & $0.27$ & $0.27$ & $0.42$ & $1.01$   & $0.06$   & $0.07$   & $0.15$   & $(0.29)$ \\
    2 & $0.13$ & $0.19$ & $0.33$ & $1.62$   & $(0.10)$ & $(0.11)$ & $(0.14)$ & *      \\
    3 & $0.33$ & $0.35$ & $0.57$ & $(1.03)$ & $0.12$   & $(0.14)$ & $0.17$   & $0.32$ \\
    4 & $0.47$ & $0.52$ & $0.81$ & *        & *        & *        & *        & *      \\
    5-8 & * & * & * & * & * & * & * & * \\
    \hline \hline
    MgO & \multicolumn{4}{c|}{Site 3} & \multicolumn{4}{c|}{Site 4} \\
    \hline
    $n$ & 13K & 32K & 60K & 106K & 13K & 32K & 60K & 106K \\
    \hline
    1 & $0.23$ & $0.26$ & $0.29$ & $0.41$ & $0.24$ & $0.26$ & $0.53$ & $1.17$ \\
    2 & $0.84$ & $0.87$ & $0.91$ & $1.04$ & $0.40$ & $0.46$ & $0.73$ & *      \\
    3 & $0.82$ & $0.86$ & $0.89$ & $1.01$ & *      & *      & *      & *      \\
    4 & $0.29$ & $0.30$ & $0.33$ & $0.57$ & *      & *      & *      & *      \\
    5-8 & * & * & * & * & * & * & * & * \\
    \hline
  \end{tabular}
  \caption{Fitted linewidths in \wn\ for all $Z_{1}Y_{n}$ excitation transitions in Sites 1, 2, 3, and 4, at 13, 32, 60, and 106 K. Peaks where a Gaussian distribution fit better than a Lorentzian based on the $R^{2}$ value of the fit are listed in parentheses. Peaks that were not visible are marked with an asterisk.}
  \label{tab:mgolinewidths}
\end{table}

@@

\begin{figure}[b]
  \centering
  \includegraphics[width=0.48\textwidth]{Site3_y1}
  \includegraphics[width=0.48\textwidth]{Site3_y2}
  \caption{Excitation peaks at $Z_1Y_1$ (left) and $Z_1Y_2$ (right) in Site 3. The intensity profile discussed in Equation \eqref{eq:1} is fitted to the $Z_{1}Y_{2}$ data.}
  \label{fig:site3peaks}
\end{figure}

@@

To understand this, we first take a step back to discuss first-order Stark shifts and inhomogeneous broadening. The standard assumption underlying the Gaussian functional form associated with inhomogeneous broadening is that ions throughout the crystal experience different random electric fields $\mathbf{F}$, which are Gaussian distributed with zero mean. The energy levels are then shifted according to first-order perturbation theory, by diagonalizing the matrix 
\begin{equation}\label{eq:2}

\end{equation}

The form of the intensity $I$ as a function of the detuning $x$ is thus \begin{equation}\label{eq:1}
  I(x) = N\int_{x'}dx'\frac{1}{x'^{2}+(\Gamma_{3}/2)^{2}}\Big(
    a_{1} D(x - \Delta_{1}) e^{-x/\Gamma_{1}} + a_{2}D(\Delta_{2}-x)e^{x/\Gamma_{2}}
  \Big) ,
\end{equation}
where $D$ is the Heaviside step function, $N$ is a normalization prefactor, $\Gamma_{1}$-$\Gamma_{3}$ are linewidth parameters, and $\Delta_{1}$-$\Delta_{2}$ are

\notetoself{In this section, I will discuss the findings of the linewidth analysis. I'm still working on it, but the goal is to fit both Gaussians and Lorentzians to the excitation peaks at $Z_{1}Y_{n}$, for all $n$, for all sites, for all temperatures. This will show when the lines are homogenously/inhomogeneously broadened, what the linewidths are, how they depend on temperature, and how things look at different sites. I'll also do a slightly more detailed discussion of the really bright, bimodal Site 3 peak. I will also include a table of the linewidths measured at each site by whichever functional form fit best, and the name of that functional form.}

\textbf{Potential for single-ion applications}: The observation of \erbium ions implanted in MgO is very exciting, because MgO has several desirable qualities as a potential host crystal. Magnesium and Oxygen both have stable isotopes with nuclear spin 0, so MgO can be made free of the nuclear spin problem that plagued YSO. This could enable substantially longer coherence times for \erbium spin qubits in MgO.

\notetoself{Another factor is our ability to stamp photonic crystals on top. Can we do this for MgO? Is there evidence we will be able to?}

\notetoself{The rest of this discussion depends a little bit on the linewidth findings. However, I'd like to discuss the strengths and weaknesses of the different sites. }

@@

% \begin{figure}[b]
%   \centering
%   \includegraphics[scale=0.97]{JinD_site1_13K}\hfill
%   \includegraphics[scale=0.97]{JinD_site1_32K}\\
%   \includegraphics[scale=0.97]{JinD_site1_60K}\hfill
%   \includegraphics[scale=0.97]{JinD_site1_106K}
%   \caption{Transition assignments for excitation-emission spectra of MgO for Site 1 at 13, 32, 60, and 106K respectively.}
%   \label{fig:mgosite1}
% \end{figure}

% \begin{figure}[b]
%   \centering
%   \includegraphics[scale=0.97]{JinD_site2_13K}\hfill
%   \includegraphics[scale=0.97]{JinD_site2_32K}\\
%   \includegraphics[scale=0.97]{JinD_site2_60K}\hfill
%   \includegraphics[scale=0.97]{JinD_site2_106K}
%   \caption{Transition assignments for excitation-emission spectra of MgO for Site 2 at 13, 32, 60, and 106K respectively.}
%   \label{fig:mgosite2}
% \end{figure}

% \begin{figure}[b]
%   \centering
%   \includegraphics[scale=0.97]{JinD_site3_13K}\hfill
%   \includegraphics[scale=0.97]{JinD_site3_32K}\\
%   \includegraphics[scale=0.97]{JinD_site3_60K}\hfill
%   \includegraphics[scale=0.97]{JinD_site3_106K}
%   \caption{Transition assignments for excitation-emission spectra of MgO for Site 3 at 13, 32, 60, and 106K respectively. Emission lines for which no level assignment can be determined are plotted in yellow.}
%   \label{fig:mgosite3}
% \end{figure}

% \begin{figure}[b]
%   \centering
%   \includegraphics[scale=0.97]{JinD_site4_13K}\hfill
%   \includegraphics[scale=0.97]{JinD_site4_32K}\\
%   \includegraphics[scale=0.97]{JinD_site4_60K}\hfill
%   \includegraphics[scale=0.97]{JinD_site4_106K}
%   \caption{Transition assignments for excitation-emission spectra of MgO for Site 4 at 13, 32, 60, and 106K respectively. No data exists for $Z_{1}Y_{1}$ at 13K, so this line is dotted in that figure.}
%   \label{fig:mgosite4}
% \end{figure}

>>>>

Unfortunately, using bulk doping methods it is impossible to grow crystals with lower concentrations of \erbium than in this previous work \cite{Dibos2017}. The Thompson lab has thus turned to ion implantation to achieve lower doses of \erbium in new potential crystal hosts.\footnote{Ion implantation is a process by which ions are accelerated into the surface of a crystal, and become lodged in the structure of the crystal. More information on this process is given in Section } Last fall, it was shown that \erbium could be successfully implanted in \TiO[]. The implanted \erbium had narrow optical linewidths, and \TiO has no nuclear spins, so this was an exciting result \cite{Phenicie2019}. However, the quality factor of photonic crystals on \TiO could never be made high enough, likely due to the poor telecom transparency of \TiO[].

In the interests of discovering better potential host crystals, a large batch of crystals was implanted with \erbium ions. For all these implanted crystals, it was necessary to determine experimentally whether \erbium was visible in the crystal, and if so, to determine its spectral properties to begin understanding how \erbium ions in this crystal would perform for single ion applications. Where are the emission and excitation lines? How bright and broad are they? What can we learn about how the \erbium ions are sitting in the crystal lattice? These are the questions that motivate this work. Before laying out our methods and results, we first provide some theoretical background about \erbium and crystals.

@@

In general, the number of crystal field parameters is infinite. However, various symmetries of the crystal can eliminate many of the possible parameters. For example, if the rare earth ion sees a crystal environment with cubic symmetry, only four crystal field parameters can be nonzero, $B_{4}^{0}$, $B_{4}^{4}$, $B_{6}^{0}$, and $B_{6}^{4}$, and only two of these are actually independent. Thus the problem is a two-parameter one. Lea, Leask and Wolf show an efficient way of calculating the crystal field matrix for diagonalization, 
\begin{equation}\label{eq:7}
  H = B_{4}(O_{4}^{0}+ 5\cdot O_{4}^{4}) + B_{6}(O_{6}^{0}-21\cdot O_{6}^{4}),
\end{equation}
where $O_{4}^{0}$, $O_{4}^{4}$, $O_{6}^{0}$ and $O_{6}^{4}$ are $2J+1$ dimensional constant matrices (for their exact form see \cite{Lea1962}). The upshot of this for our purposes is that the crystal field environment that \erbium ions experience in cubic crystals can potentially be determined exactly, enabling a complete description of the eigenstates of the \erbium optical electron. % In Section \ref{sec:mgo} we use this model to study \erbium in MgO, a crystal with cubic symmetry.

@@

\subsection{Sample preparation}
Implanted with $10^{19}$ dose \notetoself{find out units on this}. Unannealed.

\subsection{Structure}
Magnesium oxide (MgO) is a cubic lattice consisting of O$^{2-}$ ions and Mg$^{2+}$ ions held together by ionic bonding. There are three sites with cubic symmetry: the two substitutional sites have cubic symmetry, as does the interstitial site in the center of the cube. Because the crystal must remain net neutral, there might be some restructuring of the surrounding environment to compensate the introduction of the additional positive charge from the implanted \erbium ion. This may disturb the symmetries of the crystal.

\newpage 
\subsection{Energy levels}
The spectrum for MgO is fairly complicated, with at least four clearly different sites contributing to the excitation-emission spectrum. The experimentally determined energy levels for each of these sites is listed in Table \ref{tab:mgolevels}, and plotted over the spectra in Figures \ref{fig:mgosite1}, \ref{fig:mgosite2}, \ref{fig:mgosite3}, and \ref{fig:mgosite4}. These figures show excitation and emission from new $Z$ and $Y$ levels beginning to appear at higher temperatures.

Our energy level assignments are also supported by the temperature-dependent data in Figure \ref{fig:mgotemp}. These figures show the emission spectra when exciting the $Z_{1}Y_{1}$ transition of the different sites, normalized so that emission at $Y_{1}Z_{1}$ is equal to 1. The other features identified as emission from $Y_{1}$ remain the same height relative to $Y_{1}Z_{1}$. The features identified as emission from higher $Y$ levels, on the other hand, get brighter as the temperature increases, demonstrating that their thermal occupation has increased. An important assumption in this situation is that emission is slow enough that the excited state electrons thermalize fully before decaying. This assumption is supported by the fact that the relative heights of the peaks in the emission spectrum are independent of the excitation transition at which they are observed.

A few idiosyncracies are important to note for some of these sites. The $Z_{5}$ energy at Site 1 was never directly observed in the form of $Y_{1}Z_{5}$ emission like all the other $Z$ levels, and is hence listed in parenthesis. However, its location can be inferred from an emission line $Y_{3}Z_{5}$ which appears at higher temperatures. This peak can be seen emerging in Figure \ref{fig:mgosite1} and Figure \ref{fig:mgotemp}.

\begin{table}[t]
  \centering
  \begin{tabular}{| c | c | c | c | c |}
    \hline
    MgO & \multicolumn{2}{c|}{Site 1}& \multicolumn{2}{c|}{Site 2} \\
    \hline
    $n$ & $Z_{n}$ & $Y_{n}$ ($+6499$ \wn) & $Z_{n}$ & $Y_{n}$ ($+6510$ \wn) \\
    \hline 
    1 & 0.0 & 0.0 & 0.0 & 0.0 \\
    2 & 26.7 & 7.8 & 41.2 & 29.1 \\
    3 & 95.3 & 34.9 & 83.8 & 58.3 \\
    4 & 128.6 & 71.7 & 156.4 & 117.2 \\
    5 & (195.2) & 131.3 & 187.2 & * \\
    6-8 & * & * & * & * \\
    \hline \hline
    MgO & \multicolumn{2}{c|}{Site 3}& \multicolumn{2}{c|}{Site 4} \\
    \hline
    $n$ & $Z_{n}$ & $Y_{n}$ ($+6492$ \wn) & $Z_{n}$ & $Y_{n}$ ($+6557$ \wn)\\
    \hline 
    1 & 0.0 & 0.0 & 0.0 & 0.0  \\
    2 & * & (48.1) & 82.8 & 42.0 \\
    3 & * & (48.4) & 168.6 & *  \\
    4 & * & (98.6) & * & * \\
    5-8 & * & * & * & * \\
    \hline
  \end{tabular}
  \caption{Experimentally determined energy levels of MgO for all visible sites. Transitions not observed are marked with an asterisk. The level $Z_{5}$ in Site 1 is not directly observed, and thus marked with parentheses. The numbering of the $Z$ levels for Site 3 is somewhat uncertain (see the discussion in Section \ref{sec:linewidths-shapes}).}
  \label{tab:mgolevels}
\end{table}

@@

A third idiosyncracy has to do with the brightest peak in the entire spectrum, which belongs to Site 3, and falls at roughly 6540 \wn\ in excitation. This peak is bimodal, as shown in Figure \ref{fig:site3peaks}. There are two explanations for this distribution. Either these two peaks correspond to different levels in the same site, or they correspond to the same level at two sites which are very similar. To emphasize the uncertainty of the numbering of these peaks, they have been marked with parentheses in Table \ref{tab:mgolevels}.

\subsection{Crystal field modelling}
Unimplanted MgO has a cubic crystal structure. In an environment with cubic symmetry, there are only two free crystal field parameters, and thus the task of fitting an experimentally determined spectrum to theory is feasible (see Section \ref{sec:crystal-fields}). Interestingly, however, neither Site 1, Site 2, or Site 4 have energy levels consistent with a cubic crystal environment. From this, we conclude that charge compensation when \erbium replaces Mg$^{2+}$ necessitates changes to the crystal structure surrounding the \erbium which break the cubic symmetry of the environment. A likely candidate is the removal of nearby Mg$^{2+}$ ions, which can occur in line with the \erbium ion two sites away, or diagonally from the \erbium ion, or possibly farther away.

@@

\subsection{Linewidths and shapes}
\label{sec:linewidths-shapes}
PLE scans of MgO were taken at 13, 32, 60, and 106K. Taking the level assignments discussed above, the excitation peaks corresponding to excitation from $Z_{1}$ to $Y_{n}$ for all visible $Y_{n}$ were located for all four sites. For each of these peaks, two profiles were fitted, a Gaussian profile and a Lorentzian profile. When the dominant source of broadening in a spectral line is homogeneous, its profile will be Lorentzian; when the dominant broadening is inhomogeneous, the profile will in general be Gaussian. Table \ref{tab:mgolinewidths} shows the results of these fits.

\begin{table}[t]
  \centering
  \begin{tabular}{|c| c| c | c | c | c | c | c | c|}
    \hline
    MgO & \multicolumn{4}{c|}{Site 1} & \multicolumn{4}{c|}{Site 2} \\
    \hline
    $n$ & 13K & 32K & 60K & 106K & 13K & 32K & 60K & 106K \\
    \hline
    1 & $0.27$ & $0.27$ & $0.42$ & $1.01$   & $0.06$   & $0.07$   & $0.15$   & $(0.29)$ \\
    2 & $0.13$ & $0.19$ & $0.33$ & $1.62$   & $(0.10)$ & $(0.11)$ & $(0.14)$ & *      \\
    3 & $0.33$ & $0.35$ & $0.57$ & $(1.03)$ & $0.12$   & $(0.14)$ & $0.17$   & $0.32$ \\
    4 & $0.47$ & $0.52$ & $0.81$ & *        & *        & *        & *        & *      \\
    5-8 & * & * & * & * & * & * & * & * \\
    \hline \hline
    MgO & \multicolumn{4}{c|}{Site 3} & \multicolumn{4}{c|}{Site 4} \\
    \hline
    $n$ & 13K & 32K & 60K & 106K & 13K & 32K & 60K & 106K \\
    \hline
    1 & $[0.23]$ & $[0.26]$ & $0.29$ & $0.41$ & $0.24$ & $0.26$ & $0.53$ & $1.17$ \\
    2 & $[0.84]$ & $[0.87]$ & $[0.91]$ & $1.04$ & $0.40$ & $0.46$ & $0.73$ & *      \\
    3 & $[0.82]$ & $[0.86]$ & $[0.89]$ & $1.01$ & *      & *      & *      & *      \\
    4 & $0.29$ & $0.30$ & $0.33$ & $0.57$ & *      & *      & *      & *      \\
    5-8 & * & * & * & * & * & * & * & * \\
    \hline
  \end{tabular}
  \caption{Fitted linewidths in \wn\ for all $Z_{1}Y_{n}$ excitation transitions in Sites 1, 2, 3, and 4, at 13, 32, 60, and 106 K. Peaks where a Gaussian distribution fit better than a Lorentzian based on the $R^{2}$ value of the fit are listed in parentheses. Peaks with bimodal lineshapes that required further analysis are marked with square brackets. Peaks that were not visible are marked with an asterisk.}
  \label{tab:mgolinewidths}
\end{table}

@@

\begin{figure}[b]
  \centering
  \includegraphics[width=0.48\textwidth]{Site3_y1}
  \includegraphics[width=0.48\textwidth]{Site3_y2}
  \caption{Excitation peaks at $Z_1Y_1$ (left) and $Z_1Y_2$ (right) in Site 3. The intensity profile discussed in Equation \eqref{eq:1} is fitted to the $Z_{1}Y_{2}$ data. Fitted model parameters are: $a_{1} = 2.3$, $\Delta_{1} = 0.05$, $\Gamma_{1} = 0.28$, $a_{2} = 1.6$, $\Delta_{2} = -0.05$, $\Gamma_{2} = 0.28$, $\Gamma_{3} = 0.0025$, $b = 0.12$}
  \label{fig:site3peaks}
\end{figure}

@@

To understand this, we first take a step back to discuss first-order Stark shifts and inhomogeneous broadening. The standard assumption underlying the Gaussian functional form associated with inhomogeneous broadening is that ions throughout the crystal experience different random electric fields $\mathbf{F}$, which are Gaussian distributed with zero mean. The energy levels are then shifted according to first-order perturbation theory, by diagonalizing the matrix $V$ given by
\begin{equation}\label{eq:2}
  (V)_{ij} = -\mathbf{F}e\bra{\psi_{i}}\mathbf{r}\ket{\psi_{j}},
\end{equation}
where the indices $i$ and $j$ run over all eigenstates with the same energy. The key takeaway of this is that the shifts are linear in the applied field $\mathbf{F}$. Thus, since the distribution of electric fields $F$ is Gaussian, so must be the distribution of energy level shifts.

However, if the electric dipole terms $\bra{\psi_{i}}\mathbf{r}\ket{\psi_{j}}$ of the eigenstates in question are all zero, there will be no first order Stark shift. In this case, we must go to second order in perturbation theory. The second-order shift to the energy of a state $E_{i}$ is 
\begin{equation}\label{eq:8}
E_{i}^{(2)} = \sum_{j\neq i }\frac{\bra{\psi_{i}}e\mathbf{F}\cdot\mathbf{r}\ket{\psi_{j}}\bra{\psi_{j}}e\mathbf{F}\cdot\mathbf{r}\ket{\psi_{i}}}{E_{i}^{(0)}-E_{j}^{(0)}}.
\end{equation}
Now, the perturbation is proportional to the square of applied electric field. Thus, still assuming a Gaussian distribution of fields, the distribution of perturbations is exponential. Whether the distribution lies to the left or right of zero depends on the sign of the second-order Stark shift.

Understanding this, the distribution on the right side of Figure \ref{fig:site3peaks} looks a lot like two exponentials with tails facing in opposite directions. We designed an dummy intensity profile $I$ which corresponds to two exponential distributions facing in opposite directions, convolved with a Lorentzian distribution to model the contribution of the natural linewidth. $I$ as a function of the detuning $x$ is given by \begin{equation}\label{eq:1}
  I(x) = b+N\int_{x'}dx'\frac{1}{x'^{2}+(\Gamma_{3}/2)^{2}}\Big(
    a_{1} D(x - \Delta_{1}) e^{-x/\Gamma_{1}} + a_{2}D(\Delta_{2}-x)e^{x/\Gamma_{2}}
  \Big) ,
\end{equation}
where $D$ is the Heaviside step function, $N$ is a normalization prefactor, $b$ is the background intensity, $\Gamma_{1}$-$\Gamma_{3}$ are linewidth parameters, and $\Delta_{1}$-$\Delta_{2}$ are the offsets of the two exponential distributions.

Figure \ref{fig:site3peaks} shows that the agreement between this fitted model and the data is very good. We thus conclude that this feature corresponds to two different features with opposite sign second-order Stark shifts. Whether these features belong to the same site or two different sites is unclear, but the former seems more likely than the latter due to their incredibly close spacing. The same question applies to the bimodal $Z_{1}Y_{1}$ feature at Site 3. For the purposes of simplicity, the $Z_{1}Y_{1}$ features has been treated as a single line throughout this analysis. However, due to their significant separation ($\sim 0.1$ \wn), the $Z_{1}Y_{2}$ features have been treated as two separate features, referred to as $Z_{1}Y_{2}$ and $Z_{1}Y_{3}$ throughout. To emphasize the uncertainty of this choice, the energy levels reported in Table \ref{tab:mgolevels} are marked with parentheses to indicate that the numbering is uncertain.

\subsection{Potential for single-ion applications}
The observation of \erbium ions implanted in MgO is very exciting, because MgO has several desirable qualities as a potential host crystal. Magnesium and Oxygen both have stable isotopes with nuclear spin 0, so MgO can be made free of the nuclear spin problem that reduced coherence times in YSO. \erbium in MgO might therefore have substantially longer coherence times.

Another quality of MgO that made it an exciting candidate host was its high level of symmetry. A highly symmetric crystal will typically reduce the static electric dipole moment of an implanted ion, and thus its susceptibility to inhomogeneous broadening via the first-order Stark effect. Our results do indeed show very narrow inhomogeneous linewidths for several of the sites, especially Site 2, where the smallest observed linewidth was around $2$ GHz and appeared to be naturally broadened. Interestingly, the only site that was inhomogeneously broadened at 13K was Site 3, the site which we showed to have no first-order Stark shift. It is surprising that the second order Stark effect at Site 3 is more substantial than the first order Stark effect at the other sites. It is not surprising, however, that Site 3 is both the only site which fit a cubic field model and the only site where the second order Stark effect was clearly visible. A cubic crystal environment would forbid a static electric dipole moment, and thus any first order Stark shift. Our results from crystal field modelling and from the analysis of lineshapes reinforce each other in this regard.

MgO is therefore a very exciting potential host crystal. It has no nuclear spins, and very narrow inhomogeneous linewidths. It remains to be seen whether photonic crystal cavities with high quality factors can be attached to MgO. This is the next great hurdle.

@@

\begin{figure}[b]
  \centering
  \includegraphics[scale=0.97]{JinD_site1_13K}\hfill
  \includegraphics[scale=0.97]{JinD_site1_32K}\\
  \includegraphics[scale=0.97]{JinD_site1_60K}\hfill
  \includegraphics[scale=0.97]{JinD_site1_106K}
  \caption{Transition assignments for excitation-emission spectra of MgO for Site 1 at 13, 32, 60, and 106K respectively. All observed peaks are marked. Green lines correspond to excitation (emission) from $Z_{1}$ ($Y_{1}$), magenta lines correspond to origin states that thermally occupy at higher temperatures.}
  \label{fig:mgosite1}
\end{figure}

\begin{figure}[b]
  \centering
  \includegraphics[scale=0.97]{JinD_site2_13K}\hfill
  \includegraphics[scale=0.97]{JinD_site2_32K}\\
  \includegraphics[scale=0.97]{JinD_site2_60K}\hfill
  \includegraphics[scale=0.97]{JinD_site2_106K}
  \caption{Transition assignments for excitation-emission spectra of MgO for Site 2 at 13, 32, 60, and 106K respectively. All observed peaks are marked. Green lines correspond to excitation (emission) from $Z_{1}$ ($Y_{1}$), magenta lines correspond to origin states that thermally occupy at higher temperatures.}
  \label{fig:mgosite2}
\end{figure}

\begin{figure}[b]
  \centering
  \includegraphics[scale=0.97]{JinD_site3_13K}\hfill
  \includegraphics[scale=0.97]{JinD_site3_32K}\\
  \includegraphics[scale=0.97]{JinD_site3_60K}\hfill
  \includegraphics[scale=0.97]{JinD_site3_106K}
  \caption{Transition assignments for excitation-emission spectra of MgO for Site 3 at 13, 32, 60, and 106K respectively. All observed peaks are marked. Green lines correspond to excitation (emission) from $Z_{1}$ ($Y_{1}$), magenta lines correspond to origin states that thermally occupy at higher temperatures. Emission lines for which no level assignment can be determined are plotted in yellow.}
  \label{fig:mgosite3}
\end{figure}

\begin{figure}[b]
  \centering
  \includegraphics[scale=0.97]{JinD_site4_13K}\hfill
  \includegraphics[scale=0.97]{JinD_site4_32K}\\
  \includegraphics[scale=0.97]{JinD_site4_60K}\hfill
  \includegraphics[scale=0.97]{JinD_site4_106K}
  \caption{Transition assignments for excitation-emission spectra of MgO for Site 4 at 13, 32, 60, and 106K respectively. No data exists for $Z_{1}Y_{1}$ at 13K, so this line is dotted in that figure. All observed peaks are marked. Green lines correspond to excitation (emission) from $Z_{1}$ ($Y_{1}$), magenta lines correspond to origin states that thermally occupy at higher temperatures.}
  \label{fig:mgosite4}
\end{figure}
